% Abstract
% NOTE: \begin{abstract} and \end{abstract} are in main.tex

Clinical NLP pipelines detect assertion status---negation, uncertainty, family history---with $F_1 > 0.90$~\cite{gul2025beyondnegation}, yet no existing system propagates these labels through knowledge graph construction into retrieval-augmented generation.
We formalize this \emph{epistemic propagation gap} as information loss: when assertion labels are discarded during graph construction, the Shannon entropy of the epistemic state collapses to zero, upper-bounding the faithfulness of any downstream generation that conditions on the graph.
We present \epikg, a clinical knowledge graph that closes this gap with a 7-value assertion schema carried on every edge, a tri-temporal model (document, event, and ingestion time) governed by Allen's interval algebra, and an intent-aware retrieval layer that dispatches change, current-state, and historical questions to specialized graph-traversal strategies.
On ClinicalBench ($n=400$, MIMIC-IV, MedGemma~27B), we find a sharp category interaction: change-reasoning questions are \emph{impossible} without structured normalization ($0\% \to 60\%$, $+60\pp$), and family-history questions show a comparable gain ($0\% \to 57\%$, $+57\pp$), while current-state accuracy rises from $64\%$ to $76\%$ ($+12\pp$).
In aggregate, however, KG-RAG does not lift overall hard-longitudinal accuracy ($60.8\% \to 58.5\%$, $-2.3\pp$), because categories where the LLM is already strong (historical: $94\% \to 40\%$) degrade when retrieval introduces noise.
Vanilla RAG \emph{without} epistemic structure fares worse still, dropping overall accuracy by $9.0\pp$ ($71.5\% \to 62.5\%$)---assertion-blind retrieval is strictly worse than no retrieval.
Cross-model validation with Claude Opus~4.6 reveals a model-strength interaction: KG-RAG yields $-2.2\pp$ for the stronger model ($72.5\% \to 70.2\%$), confirming that structured epistemic context is most beneficial where the LLM lacks sufficient parametric clinical knowledge.
On SliceBench, KG gains scale with patient complexity ($+5.0\pp$ for patients with $\geq 15$ encounters vs.\ $+0.6\pp$ for 1--2), though overall improvement is modest ($+2.2\pp$, CI crosses zero).
Together, these results establish that knowledge graph normalization unlocks reasoning categories that are otherwise inaccessible, but the benefit is concentrated---not uniform---and interacts with both question type and model strength.
