% ============================================================
% APPENDIX
% ============================================================

\section{DR.KNOWS Benchmark Results}
\label{app:drknows}

We evaluate \epikg on the DR.KNOWS diagnostic reasoning benchmark~\cite{drknows2025}, which measures knowledge graph traversal accuracy using real PostgreSQL-backed clinical data.

\begin{table}[ht]
\centering
\small
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Score} & \textbf{\% of Baseline} \\
\midrule
Overall & 0.420 & 49.7\% \\
Path Discovery & 44.4\% & --- \\
Reasoning Accuracy & 22.2\% & --- \\
1-hop Traversal & 50.0\% & --- \\
2-hop Traversal & 25.0\% & --- \\
3-hop Traversal & 0.0\% & --- \\
\bottomrule
\end{tabular}
\caption{DR.KNOWS benchmark results. Multi-hop accuracy degrades due to PostgreSQL CTE traversal limitations versus native graph databases---an intentional architectural tradeoff for ACID compliance and transactional guarantees.}
\label{tab:drknows}
\end{table}

The 3-hop traversal failure reflects the PostgreSQL CTE approach, which requires explicit recursive query construction rather than native graph pattern matching. This is an intentional design decision: \epikg prioritizes ACID compliance and operational simplicity (single-database deployment) over raw graph traversal performance.

\section{MedQA-USMLE Baseline}
\label{app:medqa}

To establish the baseline LLM capability, we evaluated the system's language model (Claude Opus 4.5) on MedQA-USMLE~\cite{medagentbench2025} without any retrieval augmentation.

\begin{table}[ht]
\centering
\small
\begin{tabular}{lccc}
\toprule
\textbf{Model} & \textbf{Accuracy} & \textbf{Step 1} & \textbf{N} \\
\midrule
EpiKG (LLM alone) & 81.6\% & 79.9\% & 965 \\
GPT-4 (2023) & 86.7\% & --- & 1{,}273 \\
Med-PaLM 2 (2023) & 86.5\% & --- & 1{,}273 \\
Claude 3 Opus (2024) & 78.0\% & --- & 1{,}273 \\
\bottomrule
\end{tabular}
\caption{MedQA-USMLE results. Our LLM-alone baseline is competitive, confirming that performance gains from KG augmentation reflect the system's contribution rather than a stronger base model.}
\label{tab:medqa}
\end{table}

\section{Scalability Metrics}
\label{app:scalability}

\begin{table}[ht]
\centering
\small
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Documents processed & 145 \\
Patients & 85 \\
KG nodes & 3{,}100 \\
KG edges & 8{,}803 \\
Shared concept dedup ratio & $\sim$36 nodes/patient \\
1-hop traversal latency & 0.57\,ms \\
2-hop traversal latency & 0.75\,ms \\
\bottomrule
\end{tabular}
\caption{Scalability metrics at current deployment scale. Sub-millisecond traversal latency enables real-time clinical decision support.}
\label{tab:scalability}
\end{table}

\section{System Component Scale}
\label{app:scale}

\begin{table}[ht]
\centering
\small
\begin{tabular}{lr}
\toprule
\textbf{Component} & \textbf{Scale} \\
\midrule
Clinical calculators & 201 \\
Guideline sections (RAG corpus) & 1{,}202 \\
OMOP vocabulary relationships & 20M+ \\
NLP ensemble models & 3 \\
Assertion trigger patterns & 101 \\
KG edge types & 24 \\
KG node types & 13 \\
Assertion classes & 7 \\
\bottomrule
\end{tabular}
\caption{System component scale. The OMOP vocabulary provides 20M+ concept relationships for multi-hop graph traversal.}
\label{tab:system_scale}
\end{table}

\section{Human Validation Pilot}
\label{app:human}

To assess LLM-as-judge reliability, we conducted a pilot validation study with a single board-certified physician (specialty withheld for double-blind review) who scored system outputs on a 4-point scale (correct, partially correct, incorrect, clinically dangerous) without knowledge of condition labels.

\paragraph{Pilot findings ($n=5$, single reviewer).}
The pilot covered 5 ClinicalBench C1 pairs and revealed:
\begin{itemize}[nosep]
    \item The automated LLM-as-judge scoring was overly strict on uncertainty and sequence questions, rating partially correct answers as incorrect.
    \item Gold standard answers were assessed as fully correct in 3/5 cases, partially correct in 1/5, and requiring revision in 1/5.
    \item Model answers were rated partially correct or better in 3/5 cases where the auto-scorer marked them incorrect.
\end{itemize}

\paragraph{Planned full adjudication (not yet complete).}
A two-physician blinded adjudication targeting $n \geq 30$ pairs stratified across conditions (C1/C4 for ClinicalBench; B2/B3 for SliceBench) is planned, with predefined endpoints: (1)~inter-rater $\kappa$, (2)~human--LLM judge agreement rate, (3)~discordance taxonomy (false strict, false lenient, genuine disagreement).
This study is in preparation and results are not included in the current paper.

\section{Reproducibility Details}
\label{app:reproducibility}

\paragraph{Models.} ClinicalBench main ablation (Table~\ref{tab:clinicalbench_ablation}): MedGemma 27B (answering) with LLM-as-judge scoring. ClinicalBench cross-model (Tables~\ref{tab:crossmodel},~\ref{tab:crossmodel_categories}): MedGemma 27B and Claude Opus 4.6 (claude-opus-4-20250514) with deterministic keyword evaluator. SliceBench: Claude Sonnet 4.5 (claude-sonnet-4-5-20250929, answering), Claude Opus 4.6 (judging). All experiments used temperature 0 for reproducibility.

\paragraph{Evaluation.} The cross-model results use a deterministic keyword evaluator that matches temporal and assertion keywords (e.g., ``was''/``previously'' for historical, ``has''/``current'' for present) against gold-standard categories. This evaluator strips echoed evidence preambles before scoring current\_state/historical categories to prevent cross-contamination, and uses expanded uncertainty keywords including medical hedging terms (``likely'', ``probable'', ``concerning for'').

\paragraph{Bootstrap protocol.} All confidence intervals computed with $n=2{,}000$ bootstrap resamples, seed 42, bias-corrected and accelerated (BCa) method~\cite{efron1993bootstrap}.

\paragraph{Data.} All experiments used de-identified MIMIC-IV clinical notes under the PhysioNet Credentialed Health Data Use Agreement (v1.5.0).

\paragraph{Computational resources.} ClinicalBench: approximately 3.6 hours on a single machine with GPU (MedGemma inference via Ollama); Opus cross-model validation approximately 20 minutes via Anthropic API. SliceBench: approximately 2 hours using Anthropic API.

\paragraph{Evaluator evolution.} The deterministic keyword evaluator underwent refinement during development. Key changes included: (i)~switching from substring to word-boundary matching to prevent false positives (e.g., the word ``for'' matching ``former''), and (ii)~stripping echoed evidence preambles before scoring to prevent cross-contamination between categories. All numbers reported in the main text use the final evaluator version. The appendix per-category table (Table~\ref{tab:full_percategory}) has been updated to match.

\section{Assertion Category Definitions}
\label{app:assertions}

\begin{table}[ht]
\centering
\small
\begin{tabular}{lp{6.5cm}l}
\toprule
\textbf{Assertion} & \textbf{Definition} & \textbf{Example} \\
\midrule
\textsc{present} & Condition currently affirmed & ``has diabetes'' \\
\textsc{absent} & Condition explicitly negated & ``denies chest pain'' \\
\textsc{possible} & Condition suspected, not confirmed & ``possible pneumonia'' \\
\textsc{conditional} & Contingent on specific circumstances & ``if febrile, start antibiotics'' \\
\textsc{hypothetical} & Discussed as a scenario & ``would need dialysis if\ldots'' \\
\textsc{family\_history} & Attributed to a family member & ``mother had breast cancer'' \\
\textsc{historical} & Previously true, not necessarily current & ``former smoker'' \\
\bottomrule
\end{tabular}
\caption{The 7-value assertion taxonomy used in \epikg, extending the i2b2 2010 6-class standard by separating \textsc{historical} from \textsc{family\_history}.}
\label{tab:assertion_defs}
\end{table}

\section{Intent-Aware Routing Algorithm}
\label{app:routing}

The C4g condition extends epistemic KG-RAG (C4) with a question-type classifier that routes each query to a specialized retrieval strategy. The classifier is rule-based, mapping questions to one of four intent types---\textsc{Change}, \textsc{Current\_State}, \textsc{Historical}, or \textsc{Default}---using keyword patterns in the question text (e.g., ``changed'', ``new since'', ``differ between admissions'' trigger \textsc{Change}).

For \textsc{Change} queries, the retriever partitions KG edges by \texttt{hadm\_id} (hospital admission identifier) and computes set differences across admissions: concepts added (present in later admission but not earlier), removed (present in earlier but not later), and continued (present in both). For \textsc{Current\_State} queries, the retriever filters to edges with \texttt{temporality=Current} or \texttt{valid\_to=NULL}, then deduplicates by concept to return only the most recent assertion for each clinical finding. For \textsc{Historical} queries, the retriever selects edges with \texttt{temporality=Past} and augments these with admission-based inference: concepts present in an earlier admission but absent from the latest admission are inferred as ``resolved.'' All other question types fall through to the default C4 epistemic retrieval pipeline.

\section{SliceBench Tier-Stratified Results}
\label{app:slicebench_tiers}

\begin{table}[ht]
\centering
\small
\caption{SliceBench results stratified by patient complexity tier. B2$\rightarrow$B3 $\Delta$ shows the incremental KG contribution. The KG benefit concentrates in Tier~C.}
\label{tab:slicebench_tiers}
\begin{tabular}{@{}lccccccc@{}}
\toprule
& & \multicolumn{5}{c}{Score (\%)} & \\
\cmidrule(lr){3-7}
Tier & Encounters & B0 & B1 & B2 & B3 & B4 & B2$\rightarrow$B3 $\Delta$ \\
\midrule
A & 1--2 notes & 51.7 & 66.7 & 81.1 & 81.8 & 82.6 & $+0.6\pp$ \\
B & 5--10 notes & 49.5 & 79.4 & 86.4 & 87.4 & 90.1 & $+1.0\pp$ \\
C & 15+ notes & 48.4 & 74.2 & 86.5 & 91.5 & 90.3 & $+5.0\pp$ \\
\bottomrule
\end{tabular}
\end{table}

\section{ClinicalBench Full Per-Category Results}
\label{app:percategory}

\begin{table}[ht]
\centering
\small
\begin{tabular}{lcccccc}
\toprule
\textbf{Category} & \textbf{C1} & \textbf{C2} & \textbf{C3} & \textbf{C4} & \textbf{C4g} & \textbf{C5} \\
\midrule
Negation & 100.0 & 93.6 & 81.8 & 100.0 & 86.4 & 98.2 \\
Conditional & 100.0 & 90.0 & 55.0 & 70.0 & 65.0 & 75.0 \\
Uncertainty & 37.5 & 2.5 & 7.5 & 57.5 & 32.5 & 40.0 \\
Family History & 0.0 & 0.0 & 53.3 & 70.0 & 56.7 & 56.7 \\
Sequence & 80.0 & 80.0 & 80.0 & 87.5 & 90.0 & 77.5 \\
Current State & 64.0 & 54.0 & 50.0 & 54.0 & 76.0 & 58.0 \\
Duration & 100.0 & 100.0 & 83.3 & 83.3 & 46.7 & 73.3 \\
Historical & 94.0 & 78.0 & 68.0 & 60.0 & 40.0 & 68.0 \\
Change & 0.0 & 0.0 & 3.3 & 3.3 & 60.0 & 3.3 \\
\midrule
\textbf{Overall} & \textbf{71.5} & \textbf{62.5} & \textbf{59.2} & \textbf{71.5} & \textbf{66.0} & \textbf{68.2} \\
\bottomrule
\end{tabular}
\caption{Complete ClinicalBench per-category accuracy (\%, deterministic keyword evaluator) for all six conditions (C1--C5, C4g). C4g (intent-aware routing) achieves the largest gains on change questions (+60\pp vs C4) and current state (+22\pp vs C4), while C4 leads on assertion-heavy categories.}
\label{tab:full_percategory}
\end{table}

\section{Ethics Statement}
\label{app:ethics}

This work uses de-identified clinical data from MIMIC-IV~\cite{johnson2023mimiciv} under a PhysioNet Credentialed Health Data Use Agreement. No patient re-identification was attempted. All clinical questions were generated programmatically from structured data; no new clinical annotations were created. The LLM-as-judge evaluation protocol does not involve human subjects. The system is designed for clinical decision \emph{support}, not autonomous clinical decision-making, and is not intended to replace physician judgment.
