% Section 4: Benchmark Design

Existing clinical QA benchmarks evaluate factual recall (MedQA~\cite{jin2021medqa}) or agent task completion~\cite{medagentbench2025} but do not test handling of epistemic qualifiers or reasoning over longitudinal records of varying complexity.
We introduce two complementary benchmarks built on MIMIC-IV~\cite{johnson2023mimiciv}.


\subsection{ClinicalBench: Assertion-Sensitive Clinical QA}
\label{sec:bench:clinicalbench}

ClinicalBench comprises 400 gold-standard questions over 45 MIMIC-IV patients in two tasks: \textbf{Task~A} (200 questions, negation-aware retrieval---answers depend on recognizing negated, uncertain, or family-attributed conditions) and \textbf{Task~B} (200 questions, temporal reasoning---current state, sequences, durations, longitudinal changes).
Questions span 9 categories: \textit{negation}, \textit{conditional}, \textit{uncertainty}, \textit{family\_history}, \textit{sequence}, \textit{current\_state}, \textit{duration}, \textit{historical}, and \textit{change}, each with gold-standard answers from manual chart review.

Seven ablation conditions progressively add system components, plus two diagnostic baselines:

\begin{table}[t]
\centering
\small
\caption{ClinicalBench ablation conditions. C1--C4g form the primary ablation ladder; C5--C7 are diagnostic conditions.}
\label{tab:clinicalbench_conditions}
\begin{tabular}{@{}clcccp{4.2cm}@{}}
\toprule
ID & Condition & Retrieval & Assertion & Temporal & Notes \\
\midrule
C1 & LLM Alone & None & None & None & Parametric baseline \\
C2 & + Vanilla RAG & Document & None & None & Standard retrieval \\
C3 & + KG-RAG & Graph + Doc & None & None & Graph structure only \\
C4 & + Epistemic KG-RAG & Graph + Doc & Full (7-class) & Tri-temporal & v11 primary \\
C4g & + Intent-Aware KG-RAG & Graph + Doc (type-specific) & Full (7-class) & Tri-temporal & Question-type routing (Section~\ref{sec:system:intentaware}) \\
C5 & Full System & Graph + Doc + Guidelines & Full (7-class) & Tri-temporal + Calc.\ & Moved to appendix; includes calculators and guidelines \\
\midrule
C6 & Long-Context Brute Force & All docs in prompt & None & None & Tests whether token capacity alone suffices \\
C7 & Deterministic KG Only & Structured KG query & Full (7-class) & Tri-temporal & No LLM reasoning, pure KG lookup \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Endpoints.}
We define three tiers of evaluation endpoints to structure the analysis:
\begin{itemize}[nosep]
  \item \textbf{Primary endpoint:} Hard longitudinal subset (\textit{change} + \textit{current\_state} + \textit{historical}, $n=130$). The key comparisons are C4g vs.\ C1 (does KG-RAG improve over parametric reasoning?) and C4g vs.\ C6 (does structured normalization outperform brute-force long context?).
  \item \textbf{Secondary endpoint:} Full 400-question ClinicalBench, C4g vs.\ C1. This measures overall system benefit across both assertion-sensitive and temporal reasoning categories.
  \item \textbf{Diagnostic endpoints:} Per-category C1$\to$C4g accuracy deltas and C6 vs.\ C4g per category, identifying which question types benefit most from structured KG context versus raw token capacity.
\end{itemize}

\noindent\textbf{What each transition isolates.}
The C3$\to$C4 transition isolates the \emph{assertion preservation effect}: both use graph-augmented retrieval, but only C4 carries the 7-value assertion schema and tri-temporal model.
The C4$\to$C4g transition isolates the \emph{intent-aware routing effect}: both carry the full assertion and temporal metadata, but C4g additionally dispatches question-type-specific graph traversals (e.g., cross-admission partitioning for change questions, recency-weighted retrieval for current-state questions).
C6 tests whether brute-force long-context prompting---concatenating all patient documents into the prompt---eliminates the need for structured KG normalization.
C7 tests the converse: whether the knowledge graph alone, queried deterministically without LLM reasoning, suffices for clinical QA.
Together, C6 and C7 bracket the design space and justify the hybrid architecture.


\subsection{SliceBench: Complexity-Stratified Longitudinal QA}
\label{sec:bench:slicebench}

SliceBench tests the hypothesis that KG-augmented retrieval provides increasing benefit as patient record complexity grows.
We select 6 MIMIC-IV patients in three tiers: \textbf{Tier~A} (2 patients, 1--2 encounters), \textbf{Tier~B} (2 patients, 5--10 encounters), and \textbf{Tier~C} (2 patients, 15+ encounters).
Each patient receives 24 questions (categories A1--F6, where F1--F6 are hard longitudinal questions: cross-encounter medication timelines, problem list reconciliation, causal chain tracing), yielding 144 total questions.

\begin{table}[t]
\centering
\small
\caption{SliceBench conditions. B0--B4 form a monotone context progression.}
\label{tab:slicebench_conditions}
\begin{tabular}{@{}clp{7.5cm}@{}}
\toprule
ID & Condition & Context Provided to LLM \\
\midrule
B0 & LLM Alone & Question only (parametric knowledge) \\
B1 & Latest Note & Most recent clinical note \\
B2 & All Notes RAG & All notes, retrieved by relevance \\
B3 & KG-RAG & Knowledge graph context + all notes \\
B4 & Full System & KG-RAG + guidelines + calculators \\
\bottomrule
\end{tabular}
\end{table}

\noindent The critical comparison is B2$\to$B3: both have access to the same documents, but B3 additionally provides structured KG context with assertion metadata and temporal relations.

\subsection{Evaluation Protocol}
\label{sec:bench:protocol}

The two benchmarks use different evaluation strategies, chosen to match their design goals:

\paragraph{ClinicalBench} uses a \emph{deterministic keyword evaluator} that performs exact word-boundary matching of gold-standard assertion and temporal keywords against model answers.
This evaluator is fully reproducible (no stochastic LLM judge) and is used for all ClinicalBench ablation results (Table~\ref{tab:clinicalbench_ablation}), per-category breakdowns, and cross-model validation (Section~\ref{sec:results:crossmodel}).
The answering model is MedGemma 27B (primary) or Claude Opus 4.6 (cross-model validation).

\paragraph{SliceBench} uses an \emph{LLM-as-judge} protocol with separated answering and judging models to prevent self-evaluation bias~\cite{xiong2024mirage}: Claude Sonnet 4.5 generates answers and Claude Opus 4.6 judges correctness.
This allows richer semantic assessment of the open-ended longitudinal questions in SliceBench.

\paragraph{Statistical reporting.}
We report BCa bootstrap 95\% CIs ($n=2{,}000$, seed 42) on all aggregate metrics; CIs are computed at the question level (not cluster-bootstrapped by patient).
The safety score $S = 1 - \frac{1}{N} \sum_{i=1}^{N} w_i \cdot \mathbb{1}[\hat{y}_i \neq y_i]$ applies $w_i = 2.0$ for false-positive assertion errors (reporting a negated or absent condition as present) and $w_i = 1.0$ otherwise, capturing asymmetric clinical risk where acting on a falsely affirmed condition is more dangerous than missing a present one.
The asymmetric weight reflects clinical practice: acting on a falsely affirmed condition (e.g., treating a negated diagnosis) carries greater immediate risk than failing to act on a present condition, which may be caught in subsequent encounters. We treat $w=2.0$ as a reasonable default; sensitivity to this choice is a direction for future work.
