% Section 6: Discussion and Limitations
% NOTE: \section{Discussion and Limitations} and \label{sec:discussion} are in main.tex

\paragraph{Why long context underperforms.}
Brute-force long context (C6) scores 39.0\%, outperforming C1 ($+17.2\pp$) because it at least provides clinical text for the model to draw on, but it still falls far short of C4g (69.0\%) and fails completely on historical (0\%) and sequence (0\%).
C6 concatenates all notes chronologically without retrieval or selection; more sophisticated long-context strategies (hierarchical summarization, temporal segmentation) might perform better, but C6's category-level failures---consistent with the ``lost in the middle'' phenomenon~\cite{liu2024lostmiddle}---demonstrate that raw context dumping cannot substitute for structured retrieval.
Historical questions require distinguishing resolved from active conditions across admissions; sequence questions require strict temporal ordering---both are implicit in raw text but explicit in the epistemic KG.
C6 handles change detection modestly (23\%) because the relevant information appears explicitly in the text, but KG-RAG achieves $100\%$ ($+77\pp$).

\paragraph{Model-dependent KG benefit.}
Across the four models tested, KG-RAG benefit shows a consistent inverse trend: models with lower C1 baselines gain more (Table~\ref{tab:crossmodel}).
The two general-purpose models (Opus 21.8\%, GPT-OSS 20.5\%) gain $+47\pp$ and $+38\pp$ respectively; the medically fine-tuned model (MedGemma, 26.2\%) gains $+32\pp$; and the reasoning-optimized model (Qwen3.5, 37.0\%) gains least ($+21\pp$).
Despite a $17\pp$ spread at baseline, all four models converge to 57.5--69.0\% under C4g, consistent with the epistemic KG providing a knowledge floor that partially compensates for model-specific weaknesses.
With four models from distinct training lineages, the convergence pattern is consistent across all observed data points, though $n = 4$ models is insufficient for statistical inference on the correlation; additional models are needed to confirm the trend.
This has practical implications for RAG system design: the same retrieval pipeline may yield different gains depending on the downstream model, and single-model evaluation risks overstating or understating value.

\paragraph{When KG context matters.}
The complexity-dependent SliceBench effect (Tier~C: $+5.0\pp$ vs.\ Tier~A: $+0.6\pp$) suggests structured epistemic context is most valuable when: (a)~document volume overwhelms chunk-level retrieval, (b)~questions require cross-encounter synthesis where the same concept appears with different assertion states, and (c)~assertion status evolves across encounters.

The C4 ablation cleanly decomposes the C3$\to$C4g gap ($+19.0\pp$) into two components: assertions alone (C3$\to$C4: $-3.2\pp$, CI crossing zero) and intent routing (C4$\to$C4g: $+22.2\pp$, $p < 10^{-10}$).
This reveals a key architectural insight: assertion metadata does not improve aggregate accuracy without question-type-specific retrieval (CI crossing zero), and its value is unlocked only when paired with routing.
Without routing, assertions increase evidence verbosity and do not improve temporal categories (historical: $-30\pp$, sequence: $-47.5\pp$, per-category effects exploratory); with routing, the same metadata enables near-perfect change detection (C4 10\% $\to$ C4g 100\%).
The finding validates the two-component design and suggests that future work on assertion-aware clinical NLP should couple enriched metadata with structured retrieval strategies.

\paragraph{Uncertainty remains weakest.}
Uncertainty improves from 12.5\% to 50.0\% ($+37.5\pp$), but absolute accuracy remains low.
Possible and suspected conditions are linguistically subtle (``consider statin,'' ``may have CHF'') and require pragmatic inference beyond keyword matching.

\begin{sloppypar}
\paragraph{Implications beyond clinical NLP.}
The finding that metadata enrichment without retrieval alignment \emph{hurts} performance may generalize beyond clinical KGs.
Any domain where structured annotations---sentiment, stance, certainty, provenance---are extracted but not integrated into the retrieval policy faces an analogous propagation gap.
Legal reasoning requires distinguishing holdings from dicta; financial analysis must separate confirmed earnings from guidance; scientific synthesis depends on whether a result was replicated or retracted.
In each case, the design principle is the same: preserve domain-specific metadata end-to-end \emph{and} route retrieval by query intent.
The ClinicalBench ablation (C3$\to$C4$\to$C4g) provides a template for evaluating this principle in other domains.
\end{sloppypar}

\paragraph{Threats to validity and claim scope.}
\begin{sloppypar}
Our claims are bounded by dataset, evaluator, and model scope.
Data scope is limited: ClinicalBench uses 43 patients (400 questions) and SliceBench 6 patients (144 questions), all from MIMIC-IV~\cite{johnson2023mimiciv}; tier-level SliceBench effects remain hypothesis-generating ($n\!=\!2$ patients per tier), and B2$\rightarrow$B3 crosses zero (\ci{-1.5\pp}{+5.9\pp}).
Evaluator uncertainty is substantial: keyword scoring agrees with physician judgment in 47\% of the pilot, LLM-as-judge reaches 57\%, and the three evaluators produce different absolute deltas.
We therefore report all three and focus conclusions on directional agreement, not a single ``true'' effect size.
Benchmark label quality is imperfect: physician review found a 40\% error/revision rate in auto-generated gold answers (concentrated in change and negation), and pilot adjudication used a single reviewer.
Model dependence is real: gains vary by model (Table~\ref{tab:crossmodel}), so Opus-scale effects should not be assumed universal.
This study is an ablation analysis rather than a head-to-head leaderboard against existing medical RAG systems~\cite{luo2025gfmrag,jiang2025kare,lu2025doctorrag}; C2 is used as a proxy baseline for document-level retrieval on our longitudinal task.
Runtime variance from quantized local inference (e.g., MedGemma $\pm10\pp$ across runs) and single-site provenance further limit external validity.
Finally, this work does \emph{not} establish clinical deployment readiness: it evaluates QA behavior under controlled benchmarks, not prospective patient outcomes.
\end{sloppypar}

\begin{sloppypar}
\paragraph{Conclusion.}
The epistemic propagation gap is a real and measurable phenomenon: clinical NLP systems that extract assertion status but discard it before retrieval systematically misrepresent patient state.
The key finding is that closing this gap requires two components: assertion metadata alone does not improve accuracy (C3$\to$C4: $-3.2\pp$, CI crossing zero), but intent-aware routing unlocks its value (C4$\to$C4g: $+22.2\pp$, $p < 10^{-10}$).
This holds across four models from distinct training lineages, with a convergence pattern in which all four cluster at 57.5--69.0\% under C4g despite divergent baselines---consistent with the epistemic KG providing a knowledge floor.
The design principle---preserve epistemic metadata, route retrieval by question intent, evaluate by category$\times$condition interaction---is falsifiable, domain-agnostic, and immediately testable in any RAG system operating over annotated structured data.
\end{sloppypar}
