% Section 1: Introduction
% NOTE: \section{Introduction} and \label{sec:intro} are in main.tex

A physician documents ``patient denies chest pain, sister had MI at 45, will consider statin if lipids remain elevated''---a single sentence encoding four epistemic states: a negated symptom, a family-history event, a hypothetical action, and an implicit present condition.
Clinical NLP can detect these assertions accurately~\cite{uzuner2011i2b2assertion, gul2025beyondnegation}, yet when a RAG system ingests the same note, assertion context is flattened: negation is lost, family attribution collapses onto the patient, and downstream reasoning may conclude the patient \emph{has} chest pain and \emph{had} an MI.
We call this the \emph{epistemic propagation gap}---the systematic loss of assertion status as clinical information moves from extraction through storage to retrieval.

The gap persists because, to our knowledge, no system connects assertion detection to knowledge representation.
The i2b2 2010 six-class taxonomy~\cite{uzuner2011i2b2assertion} remains the standard, yet we are unaware of a subsequent system that propagates these labels beyond extraction.
OMOP excludes negated conditions from \texttt{CONDITION\_OCCURRENCE}~\cite{omop2024}; FHIR provides \texttt{verificationStatus} for conditions only.
Graph-augmented RAG systems---GraphRAG~\cite{edge2024graphrag}, GFM-RAG~\cite{luo2025gfmrag}, KARE~\cite{jiang2025kare}, Medical-Graph-RAG~\cite{wu2025medicalgraphrag}---build KGs that discard the metadata distinguishing ``patient has diabetes'' from ``rule out diabetes.''
Two KG surveys~\cite{alkhatib2024pckgsurvey,abusalih2023healthcarekg} do not identify assertion preservation as a concern.
A parallel \emph{temporal integration gap} exists: clinical events span three time dimensions (valid time, transaction time, NLP-asserted temporality), yet existing systems model at most two~\cite{huang2024medtkg,ramirez2025graphiti}, and Allen's interval algebra~\cite{allen1983temporal} has been formalized only as annotation ontologies~\cite{li2020teo} or modal operators~\cite{ong2023tcl}, never as first-class KG edge attributes.

We present \epikg, a clinical KG system that closes both gaps by preserving epistemic status end-to-end from NLP extraction through OMOP mapping, fact construction, KG materialization, and graph-augmented retrieval.
Our contributions:

\begin{enumerate}[leftmargin=*, topsep=4pt, itemsep=2pt]
  \item \textbf{Formal epistemic preservation invariant.} We define the assertion preservation invariant---that the epistemic status assigned at extraction must be recoverable at every downstream stage---and provide an information-theoretic characterization: assertion-blind pipelines collapse assertion entropy to zero, incurring quantifiable information loss that bounds downstream faithfulness (Section~\ref{sec:formal_invariant}).

  \item \textbf{End-to-end assertion preservation.} Subordinate to the invariant, we implement a seven-value assertion schema extending i2b2~\cite{uzuner2011i2b2assertion}, carried intact from NLP extraction through OMOP mapping, fact construction, KG materialization, and RAG context serialization---to our knowledge, no prior system maintains this chain.

  \item \textbf{Tri-temporal KG with Allen's interval algebra.} Edges carry valid time, transaction time, and NLP-asserted temporality alongside nine Allen's relations as first-class attributes, enabling temporal reasoning over longitudinal records.

  \item \textbf{Shared concept node architecture.} Global OMOP concept deduplication with patient-specific edges enables both per-patient reasoning and cross-patient cohort queries without node duplication.

  \item \textbf{Seven-condition falsifiable design space.} We evaluate seven retrieval conditions---C1 (no retrieval) through C5 (full system with calculators and guidelines)---plus bookend baselines C6 (brute-force full-context) and C7 (structured-only, no LLM). The bookend conditions make the design space falsifiable: C6 upper-bounds what unstructured access can achieve, and C7 isolates the KG's contribution independent of LLM reasoning.

  \item \textbf{Assertion-sensitive benchmarks with physician adjudication.} ClinicalBench (400 questions, 9 epistemic categories) targets the epistemic propagation gap, with hard longitudinal questions (change, cross-admission, temporal) as the primary endpoint. Blinded physician adjudication validates automated scoring on a stratified sample.
\end{enumerate}

The central research question is: \emph{under what conditions does structured epistemic context help, hurt, or break even compared to unstructured retrieval or no retrieval at all?} The answer is nuanced. KG-RAG yields large gains on assertion-sensitive and longitudinal categories where parametric knowledge is insufficient---but it can hurt on categories where the LLM's parametric knowledge already suffices and added context introduces noise. The magnitude and even direction of the effect depend on model strength, producing a model$\times$condition interaction that aggregate accuracy alone obscures.

On the primary endpoint of hard longitudinal questions, \epikg's assertion-aware KG-RAG produces substantial gains over no retrieval, while vanilla RAG \emph{degrades} performance ($-9.0\pp$)---retrieving documents without epistemic structure actively harms faithfulness. Category-level analysis reveals the mechanism: $+20\pp$ on uncertainty, $+70\pp$ on family history, with the benefit concentrating where assertion status is decision-relevant. The effect interacts with model strength: stronger models extract more value from structured context, while weaker models are more susceptible to context noise, a finding with direct implications for clinical deployment.
